% =============================================================================
% Capítulo 7 — Referencias
% Alineado con propuesta final + referencias adicionales usadas en informe final
% =============================================================================

\newpage
\section{Referencias}

\begin{list}{}{%
    \setlength{\leftmargin}{1.27cm}
    \setlength{\itemindent}{-1.27cm}
    \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
    \setlength{\topsep}{0pt}
    \setlength{\listparindent}{0pt}
    \setlength{\labelwidth}{0pt}
    \setlength{\labelsep}{0pt}
}
\RaggedRight
\sloppy
\emergencystretch=3em

\item Akenine-Möller, T., Haines, E., \& Hoffman, N. (2018). \textit{Real-time rendering} (4th ed.). CRC Press.

\item Bangor, A., Kortum, P. T., \& Miller, J. T. (2008). An empirical evaluation of the system usability scale. \textit{International Journal of Human-Computer Interaction}, \textit{24}(6), 574--594. \url{https://doi.org/10.1080/10447310802205776}

\item Bangor, A., Kortum, P. T., \& Miller, J. T. (2009). Determining what individual SUS scores mean: Adding an adjective rating scale. \textit{Journal of Usability Studies}, \textit{4}(3), 114--123.

\item Babylon.js. (s.\,f.). \textit{Engine specifications}. \url{https://www.babylonjs.com/specifications/}

\item Bartlett, K. A., \& Dorribo Camba, J. (2023). The role of a graphical interpretation factor in the assessment of spatial visualization: A critical analysis. \textit{Spatial Cognition \& Computation}, \textit{23}(1), 1--30. \url{https://doi.org/10.1080/13875868.2021.2019260}

\item Blender Bros. (2022). \textit{The Blender Bros hard surface game asset course 2.0}. \url{https://www.blenderbros.com/gameasset}

\item Bowman, D. A., Kruijff, E., LaViola, J. J., Jr., \& Poupyrev, I. (2004). \textit{3D user interfaces: Theory and practice}. Addison-Wesley.

\item Brooke, J. (1996). SUS: A ``quick and dirty'' usability scale. En P. W. Jordan, B. Thomas, I. L. McClelland, \& B. Weerdmeester (Eds.), \textit{Usability evaluation in industry} (pp. 189--194). Taylor \& Francis.

\item Burley, B. (2012). Physically-based shading at Disney. \textit{ACM SIGGRAPH 2012 course notes: Practical physically based shading in film and game production}. ACM.

\item Cabello, R. [mrdoob]. (2024). \textit{three.js -- JavaScript 3D library} [Repositorio de GitHub]. GitHub. \url{https://github.com/mrdoob/three.js}

\item Cook, R. L., \& Torrance, K. E. (1982). A reflectance model for computer graphics. \textit{ACM Transactions on Graphics}, \textit{1}(1), 7--24. \url{https://doi.org/10.1145/357290.357293}

\item Cowan, N. (2001). The magical number 4 in short-term memory: A reconsideration of mental storage capacity. \textit{Behavioral and Brain Sciences}, \textit{24}(1), 87--114.

\item Darken, R. P., \& Sibert, J. L. (1996). Wayfinding strategies and behaviours in large virtual worlds. \textit{Proceedings of the SIGCHI Conference on Human Factors in Computing Systems}, 142--149. \url{https://doi.org/10.1145/238386.238459}

\item David, I., Shao, G., Tilbury, D., Gomes, C., \& Zarkhout, B. (2024). \textit{Interoperability of digital twins: Challenges, success factors, and future research directions}. National Institute of Standards and Technology. \url{https://www.nist.gov/publications/interoperability-digital-twins-challenges-success-factors-and-future-research}

\item Digital Twin Consortium. (2020, 3 de diciembre). \textit{Digital Twin Consortium defines digital twin}. \url{https://www.digitaltwinconsortium.org/2020/12/digital-twin-consortium-defines-digital-twin/}

\item Ericsson, K. A., \& Simon, H. A. (1993). \textit{Protocol analysis: Verbal reports as data} (Revised ed.). MIT Press.

\item Epic Games. (s.\,f.-a). \textit{Unreal Engine Pixel Streaming Reference}. Epic Developer Community. \url{https://dev.epicgames.com/documentation/en-us/unreal-engine/unreal-engine-pixel-streaming-reference?application_version=5.6}

\item Epic Games. (s.\,f.-b). \textit{Pixel Streaming Infrastructure}. Epic Developer Community. \url{https://dev.epicgames.com/documentation/en-us/unreal-engine/pixel-streaming-infrastructure?application_version=5.6}

\item Fransson, E., Hermansson, J., \& Hu, Y. (2024). A comparison of performance on WebGPU and WebGL in the Godot game engine. En \textit{2024 IEEE Gaming, Entertainment, and Media Conference (GEM)} (pp. 1--6). IEEE. \url{https://doi.org/10.1109/GEM61861.2024.10585437}

\item Fuller, A., Fan, Z., Day, C., \& Barlow, C. (2020). Digital twin: Enabling technologies, challenges and open research. \textit{IEEE Access}, \textit{8}, 108952--108971. \url{https://doi.org/10.1109/ACCESS.2020.2998358}

\item Hart, S. G. (2006). NASA-task load index (NASA-TLX): 20 years later. \textit{Proceedings of the Human Factors and Ergonomics Society Annual Meeting}, \textit{50}(9), 904--908. \url{https://doi.org/10.1177/154193120605000909}

\item Hart, S. G., \& Staveland, L. E. (1988). Development of NASA-TLX (Task Load Index): Results of empirical and theoretical research. En P. A. Hancock \& N. Meshkati (Eds.), \textit{Human mental workload} (pp. 139--183). North-Holland.

\item Hegarty, M., \& Waller, D. (2004). Spatial abilities at different scales: Individual differences in aptitude-test performance and spatial-layout learning. \textit{Intelligence}, \textit{32}(2), 151--176.

\item Hevner, A. R., March, S. T., Park, J., \& Ram, S. (2004). Design science in information systems research. \textit{MIS Quarterly}, \textit{28}(1), 75--105.

\item Holybro. (s.\,f.-a). \textit{PX4 development kit -- X500v2}. Holybro Docs. \url{https://docs.holybro.com/drone-development-kit/px4-development-kit-x500v2}

\item Holybro. (s.\,f.-b). \textit{Spare parts -- X500 V2 kit}. Holybro Store. \url{https://holybro.com/products/spare-parts-x500-v2-kit}

\item Holybro. (s.\,f.-c). \textit{Holybro X500 V2 frame kit assembly guide} [Guía del fabricante]. \url{https://cdn.shopify.com/s/files/1/0604/5905/7341/files/Holybro_X500_V2_Frame_Kit_Assembly_Guide_en.pdf?v=1646987799}

\item Hutchins, E. (1995). \textit{Cognition in the wild}. MIT Press.

\item International Organization for Standardization. (2021). \textit{ISO 23247-1:2021 Automation systems and integration -- Digital twin framework for manufacturing -- Part 1: Overview and general principles}. ISO. \url{https://www.iso.org/standard/75066.html}

\item Jones, D., Snider, C., Nassehi, A., Yon, J., \& Hicks, B. (2020). Characterising the digital twin: A systematic literature review. \textit{CIRP Journal of Manufacturing Science and Technology}, \textit{29}, 36--52. \url{https://doi.org/10.1016/j.cirpj.2020.02.002}

\item Kajiya, J. T. (1986). The rendering equation. \textit{ACM SIGGRAPH Computer Graphics}, \textit{20}(4), 143--150. \url{https://doi.org/10.1145/15886.15902}

\item Karis, B. (2013). Real shading in Unreal Engine 4. En B. Burley (Ed.), \textit{SIGGRAPH 2013 course: Physically based shading in theory and practice}. ACM.

\item Khronos Group. (2011). \textit{WebGL Specification 1.0}. \url{https://www.khronos.org/registry/webgl/specs/latest/1.0/}

\item Khronos Group. (2024). \textit{glTF -- Runtime 3D asset delivery}. \url{https://www.khronos.org/gltf/}

\item Kritzinger, W., Karner, M., Traar, G., Henjes, J., \& Sihn, W. (2018). Digital twin in manufacturing: A categorical literature review and classification. \textit{IFAC-PapersOnLine}, \textit{51}(11), 1016--1022. \url{https://doi.org/10.1016/j.ifacol.2018.08.474}

\item Köhler, W. (1929). \textit{Gestalt psychology}. Liveright.

\item Lazar, J., Feng, J. H., \& Hochheiser, H. (2017). \textit{Research methods in human-computer interaction} (2nd ed.). Morgan Kaufmann.

\item Lin, S.-W., Watson, K., Shao, G., Stojanovic, L., \& Zarkout, B. (2023). \textit{Digital twin core conceptual models and services}. National Institute of Standards and Technology. \url{https://www.nist.gov/publications/digital-twin-core-conceptual-models-and-services}

\item Miller, G. A. (1956). The magical number seven, plus or minus two: Some limits on our capacity for processing information. \textit{Psychological Review}, \textit{63}(2), 81--97.

\item Marmoset. (s.\,f.). \textit{Marmoset Viewer (.mview)}. Marmoset Toolbag Documentation. \url{https://docs.marmoset.co/docs/marmoset-viewer/}

\item Nielsen, J. (1994). \textit{Usability engineering}. Morgan Kaufmann.

\item Nielsen, J. (2000). Why you only need to test with 5 users. Nielsen Norman Group. \url{https://www.nngroup.com/articles/why-you-only-need-to-test-with-5-users/}

\item Norman, D. A. (2013). \textit{The design of everyday things: Revised and expanded edition}. Basic Books.

\item Open Cascade. (s.\,f.). \textit{glTF model assets}. OCCT3D Documentation. \url{https://occt3d.com/doc/occviewer/3.0.354/documents/Guide.Model_types.gltf.html}

\item Paivio, A. (1986). \textit{Mental representations: A dual coding approach}. Oxford University Press.

\item Peffers, K., Tuunanen, T., Rothenberger, M. A., \& Chatterjee, S. (2007). A design science research methodology for information systems research. \textit{Journal of Management Information Systems}, \textit{24}(3), 45--77. \url{https://doi.org/10.2753/MIS0742-1222240302}

\item PlayCanvas. (s.\,f.-a). \textit{Editor}. \url{https://playcanvas.com/products/editor}

\item PlayCanvas. (s.\,f.-b). \textit{Engine}. \url{https://playcanvas.com/products/engine}

\item PlayCanvas. (s.\,f.-c). \textit{Scripting}. \url{https://developer.playcanvas.com/user-manual/scripting/}

\item PX4 Development Team. (2026). \textit{Simulation}. PX4 User Guide. \url{https://docs.px4.io/main/en/simulation/}

\item Sauro, J., \& Lewis, J. R. (2016). \textit{Quantifying the user experience: Practical statistics for user research} (2nd ed.). Morgan Kaufmann.

\item Schlick, C. (1994). An inexpensive BRDF model for physically-based rendering. \textit{Computer Graphics Forum}, \textit{13}(3), 233--246.

\item Sharma, A., Kosasih, E., Zhang, J., Brintrup, A., \& Calinescu, A. (2022). Digital twins: State of the art theory and practice, challenges, and open research questions. \textit{Journal of Industrial Information Integration}, \textit{30}, 100383. \url{https://doi.org/10.1016/j.jii.2022.100383}

\item Shepard, R. N., \& Metzler, J. (1971). Mental rotation of three-dimensional objects. \textit{Science}, \textit{171}(3972), 701--703.

\item Sketchfab. (s.\,f.). \textit{Viewer API}. \url{https://sketchfab.com/developers/viewer}

\item Spline. (s.\,f.-a). \textit{Play mode}. \url{https://docs.spline.design/basics/play-mode}

\item Spline. (s.\,f.-b). \textit{Exporting as Spline Viewer}. \url{https://docs.spline.design/doc/exporting-as-spline-viewer/docoQp6On9ht}

\item Spline. (s.\,f.-c). \textit{Exporting as glTF/GLB}. \url{https://docs.spline.design/exporting-your-scene/files/exporting-as-gtlf-glb}

\item Sweller, J. (1988). Cognitive load during problem solving: Effects on learning. \textit{Cognitive Science}, \textit{12}(2), 257--285.

\item Sweller, J., van Merriënboer, J. J. G., \& Paas, F. (2019). Cognitive architecture and instructional design: 20 years later. \textit{Educational Psychology Review}, \textit{31}(2), 261--292. \url{https://doi.org/10.1007/s10648-019-09465-5}

\item Tao, F., Zhang, H., Liu, A., \& Nee, A. Y. C. (2019). Digital twin in industry: State-of-the-art. \textit{IEEE Transactions on Industrial Informatics}, \textit{15}(4), 2405--2415. \url{https://doi.org/10.1109/TII.2018.2873186}

\item Unity Technologies. (2024a). \textit{Memory in WebGL}. Unity Documentation. \url{https://docs.unity3d.com/Manual/webgl-memory.html}

\item Unity Technologies. (2024b). \textit{Universal Render Pipeline overview}. Unity Documentation. \url{https://docs.unity3d.com/Packages/com.unity.render-pipelines.universal@17.0/manual/}

\item Unity Technologies. (2024c). \textit{WebGL: Interacting with browser scripting}. Unity Documentation. \url{https://docs.unity3d.com/Manual/webgl-interactingwithbrowserscripting.html}

\item Unity Technologies. (2024d). \textit{UI Toolkit}. Unity Documentation. \url{https://docs.unity3d.com/Manual/UIElements.html}

\item Unity Technologies. (s.\,f.). \textit{Web browser compatibility}. Unity Documentation. \url{https://docs.unity3d.com/6000.0/Manual/webgl-browsercompatibility.html}

\item Walter, B., Marschner, S. R., Li, H., \& Torrance, K. E. (2007). Microfacet models for refraction through rough surfaces. \textit{Proceedings of the 18th Eurographics Conference on Rendering Techniques}, 195--206.

\item Ware, C. (2012). \textit{Information visualization: Perception for design} (3rd ed.). Morgan Kaufmann.

\item WebAssembly Community Group. (s.\,f.). \textit{WebAssembly core specification}. \url{https://webassembly.github.io/spec/core/}

\item Wertheimer, M. (1923). Untersuchungen zur Lehre von der Gestalt II. \textit{Psychologische Forschung}, \textit{4}(1), 301--350.

\item World Wide Web Consortium. (2023). \textit{Web of Things (WoT) Architecture 1.1}. W3C Recommendation. \url{https://www.w3.org/TR/wot-architecture11/}

\item Yu, G., Liu, C., Fang, T., Jia, J., Lin, E., He, Y., Fu, S., Wang, L., Wei, L., \& Huang, Q. (2023). A survey of real-time rendering on Web3D application. \textit{Virtual Reality \& Intelligent Hardware}, \textit{5}(5), 379--394. \url{https://doi.org/10.1016/j.vrih.2022.04.002}

\end{list}
